Approximation as Differentiation

### 2025-05-24 main/introduction.tex
weiser81.

Provenance tracking aims to answer questions like ``Where did these data come from?'', ``Why are these data in the output?'' and ``How were these data computed?''. Provenance tracking has a wide range of applications, from debugging and program comprehension to reproducibility and transparency in scientific workflows. In order to audit any computational process, we need robust and well-founded notions of provenance to track how data are used, how data are computed, and what data influenced a particular outcome. \emph{Program slicing} is a collection of techniques for provenance tracking that attempts to take a run of a program and areas of interest in the output, and turn them into the subset of the input and the program that were responsible for generating those specific outputs.

@@

Perera and collaborators introduced the idea of {\em Galois Program Slicing} as a particular conception of program slicing for provenance, described in several publications~\cite{perera12a,perera16d,ricciotti17}. Galois Program Slicing (hereafter simply {\emph \GPS}) forms the basis of the open source data visualisation tool \href{https://f.luid.org/}{Fluid} that allows interactive exploration of programmatically generated visualisations.

>>>>

Provenance tracking aims to answer questions like ``Where did these data come from?'', ``Why are these data in the output?'' and ``How were these data computed?''. Provenance tracking has a wide range of applications, from debugging and program comprehension to improving reproducibility and transparency in scientific workflows. In order to audit any computational process, we need robust and well-founded notions of provenance to track how data are used, how data are computed, and what data influenced a particular outcome. \emph{Program slicing}, first proposed by~\citet{weiser81}, is a collection of techniques for provenance tracking that attempts to take a run of a program and areas of interest in the output, and turn them into the subset of the input and the program that were responsible for generating those specific outputs.

@@

Perera and collaborators introduced the idea of {\em Galois program slicing} as a particular conception of program slicing for provenance, described in several publications~\cite{perera12a,perera16d,ricciotti17}. Galois program slicing (hereafter simply {\emph \GPS}) forms the basis of the open source data visualisation tool \href{https://f.luid.org/}{Fluid} that allows interactive exploration of programmatically generated visualisations.
